\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-12-12 proof-sgd.tex
Working on the proof of the bloody sgd theorem

This proof requires somewhat more techinicality than the first part of the
theorem, including a brief detour into local asymptotic normality theory,
regular estimators, and quadratic-mean
differentiability~\cite[cf.]{VanDerVaart98}.
We begin by recalling some of these concepts.
\begin{definition}
  A sequence of estimators $T_n$ for a parameter $\theta$ in the parametric
  family $\{P_\theta\}_{\theta \in \Theta}$ is \emph{regular at $\theta$} if
  there exists a distribution $Q$ such that for all bounded sequence
  $h_n$,
  \begin{equation*}
    \sqrt{n}(T_n - (\theta + h_n / \sqrt{n}))
    \mathop{\cd}_{P_{\theta + h_n/\sqrt{n}}}
    Q.
  \end{equation*}
\end{definition}
\begin{definition}
  Let $\{P_\theta\}_{\theta \in \Theta}$ have densities $p_\theta$
  with respect to a base measure $\mu$. The family
  is \emph{quadratic mean differentiable} (QMD) at
  $\theta$ with score $\score_\theta$ if
  \begin{equation}
    \label{eqn:qmd}
    \int \left(\sqrt{p_{\theta + h}} - \sqrt{p_\theta}
    - \half h^\top \score_\theta\sqrt{p_\theta} \right)^2 d\mu
    = o(\norm{h}^2)
  \end{equation}
  as $h \to 0$.
\end{definition}
\begin{definition}
  A family of distributions $\{P_\theta\}_{\theta \in \Theta}$ is
  \emph{locally asymptotically normal with information matrix $I_\theta$} at
  $\theta$ if there exists a sequence of random vectors $Z_n$ such that for
  all $h_n \to h$,
  \begin{equation*}
    \sum_{i = 1}^n \log \frac{dP_{\theta + h_n/\sqrt{n}}}{dP_\theta}
    (X_1, \ldots, X_n)
    = h^\top Z_n - \half h^\top I_\theta h + o_P(1)
  \end{equation*}
  where $Z_n \cd \normal(0, I_\theta)$ under $P_\theta$, where
  $X_i \simiid P_\theta$.
\end{definition}

These three definitions are linked in our case by a few important results.
First~\cite[Theorem 7.2]{VanDerVaart98},
if $P_\theta$ is QMD at $\theta$, then
it is locally asymptotically normal with
$Z_n = \frac{1}{\sqrt{n}} \sum_{i = 1}^n \score_\theta(X_i)$ and information
matrix $I_\theta = \E_\theta[\score_\theta \score_\theta^\top]$. Moreover,
in any locally asymptotically normal family at $\theta$,
if $T_n$ is a regular estimator at $\theta$ with limiting distribution $Q$,
then for any bounded, symmetric, quasi-convex loss $L$ and $c < \infty$,
\begin{equation*}
  \limsup_n \sup_{\norm{h} \le c}
  \E_{P_{\theta + h/\sqrt{n}}}
  \left[L(\sqrt{n}(T_n - \theta - h / \sqrt{n}))\right]
  = \E[L(W)] ~~ \mbox{for~}W \sim Q
\end{equation*}
(see Beran~\cite{beran1995role}, Eq.~(4.2)).

Thus, we show three results: first, that $\bar{\theta}_n$ is regular,
second, that the family $P_\theta$ is QMD, and finally, combining the two
results.

@@

This is essentially an immediate consequence of Le Cam's third
lemma~\cite[Example 6.7]{VanDerVaart98}.
We recall~\cite[Thm.~7.2]{VanDerVaart98} that if
a family $\{P_\theta\}_{\theta \in \Theta}$ is quadratic mean differentiable
at $\theta$ with score $\score_\theta$, then for any sequence
$h_n \to h$, if $Z_i \simiid P_\theta$ then
\begin{equation}
  \label{eqn:lan-from-qmd}
  \log \frac{dP_{\theta + h/\sqrt{n}}^n}{dP_\theta} (Z_1, \ldots, Z_n)
  = \frac{h}{\sqrt{n}} \sum_{i = 1}^n \score_\theta(Z_i)
  - \half h^2 I_\theta
  + o_{P_\theta,n}(1)
\end{equation}
where $I_\theta = \E_{P_\theta}[\score_\theta^2]$ is the Fisher information,
which necessarily exists.

We thus show that under the conditions we have assumed,
we have quadratic mean differentiability.
\begin{lemma}[Extension of \cite{VanDerVaart98}, Lemma 7.6]
  \label{lemma:qmd}
  Let $p_\theta$ be a density with respect to $\mu$,
  and assume that $\theta \mapsto s_\theta(x) \defeq \sqrt{p_\theta(x)}$
  is absolutely continuous for all $x$. Let
  $\dot{p}_\theta(x) = \nabla_\theta p_\theta(x)$ (when it exists),
  and assume that
  \begin{equation*}
    \mu(\{x : \dot{p}_\theta(x) ~ \mbox{fails~to~exist}\}) = 0.
  \end{equation*}
  Assume that $I_\theta \defeq \E_{P_\theta}[\dot{p}_\theta
    \dot{p}_\theta^\top / p_\theta^2]$ is continuous at $\theta_0$. Then
  $P_\theta$ is QMD~\eqref{eqn:qmd} at $\theta = \theta_0$ with
  $\score_\theta = \dot{p}_\theta / p_\theta$.
\end{lemma}

Deferring the proof of Lemma~\ref{lemma:qmd} temporarily,
we note that the regularity result~\eqref{eq:normal_expansion_lem}
gives
\begin{equation*}
  \sqrt{n} \bar{\Delta}_n = -\frac{1}{\sqrt{n}}
  \sum_{i = 1}^n  \frac{\varphi(Z_i)}{\psi'(0)} + o_{P,n}(1).
\end{equation*}
The conditions in
Corollary~\ref{corollary:normal-expansion}\eqref{item:apply-le-cam} imply
that the Fisher information $I_h =
\E_h[\score_h(Z)^2]$ exists and is continuous for
$\score_h(z) = \frac{p'(z - h)}{f(z - h)}$,
and the asymptotic expansion~\eqref{eqn:lan-from-qmd}
with the preceding display give the joint convergence
\begin{equation*}
  \left(\sqrt{n} \bar{\Delta}_n, \log \frac{P^n_{h_n/\sqrt{n}}}{P^n_0}(Z_1^n)
  \right)
  \cd \normal \left( \left(0,-\frac{h^2}{2} I_0 \right),
  \begin{pmatrix}
    \frac{\chi(0)}{\psi'(0)^2} & \frac{-h}{ \psi'(0)}
    \E_p[\varphi(Z) \score_0(Z)] \\
    \frac{-h}{ \psi'(0)} \E_p[\varphi(Z) \score_0(Z)]
    & h^2 I_0
  \end{pmatrix}  \right).
\end{equation*}
Le Cam's third lemma \cite[Exm. 6.7]{VanDerVaart98} then implies the
convergence
\begin{equation*}
  \sqrt{n}\bar{\Delta}_n
  \mathop{\cd}_{P_{h_n/\sqrt{n}}^n}
  \Ncal\left(\frac{-h}{ \psi'(0)} \E_f[\varphi(Z) \score(Z)],
  \frac{\chi(0)}{\psi'(0)^2} \right)
\end{equation*}
under the alternatives $P_{h_n/\sqrt{n}}$, which gives
Corollary~\ref{corollary:normal-expansion}\eqref{item:apply-le-cam}.

@@

\subsubsection*{Proof of Lemma~\ref{lemma:qmd}}

>>>>

This proof requires somewhat more techinicality than the first part of the
theorem, including a brief detour into local asymptotic normality theory,
regular estimators, and quadratic-mean
differentiability~\cite[cf.]{VanDerVaart98}.
We begin by recalling some of these concepts.
\begin{definition}
  \label{definition:regular-estimator}
  A sequence of estimators $T_n$ for a parameter $\theta$ in the parametric
  family $\{P_\theta\}_{\theta \in \Theta}$ is \emph{regular at $\theta$} if
  there exists a distribution $Q$ such that for all bounded sequence
  $h_n$,
  \begin{equation*}
    \sqrt{n}(T_n - (\theta + h_n / \sqrt{n}))
    \mathop{\cd}_{P_{\theta + h_n/\sqrt{n}}}
    Q.
  \end{equation*}
\end{definition}
\begin{definition}
  \label{definition:qmd}
  Let $\{P_\theta\}_{\theta \in \Theta}$ have densities $p_\theta$
  with respect to a base measure $\mu$. The family
  is \emph{quadratic mean differentiable} (QMD) at
  $\theta$ with score $\score_\theta$ if
  \begin{equation}
    \label{eqn:qmd}
    \int \left(\sqrt{p_{\theta + h}} - \sqrt{p_\theta}
    - \half h^\top \score_\theta\sqrt{p_\theta} \right)^2 d\mu
    = o(\norm{h}^2)
  \end{equation}
  as $h \to 0$.
\end{definition}
\begin{definition}
  \label{definition:lan}
  A family of distributions $\{P_\theta\}_{\theta \in \Theta}$ is
  \emph{locally asymptotically normal with information matrix $I_\theta$} (LAN)
  at
  $\theta$ if there exists a sequence of random vectors $Z_n$ such that for
  all $h_n \to h$,
  \begin{equation*}
    \sum_{i = 1}^n \log \frac{dP_{\theta + h_n/\sqrt{n}}}{dP_\theta}
    (X_1, \ldots, X_n)
    = h^\top Z_n - \half h^\top I_\theta h + o_P(1)
  \end{equation*}
  where $Z_n \cd \normal(0, I_\theta)$ under $P_\theta$, where
  $X_i \simiid P_\theta$.
\end{definition}

These three definitions are linked in our case by a few important results.
First~\cite[Theorem 7.2]{VanDerVaart98}, if $\{P_\theta\}$ is QMD
(Def.~\ref{definition:qmd}) at the point $\theta$, then it is locally
asymptotically normal with $Z_n = \frac{1}{\sqrt{n}} \sum_{i = 1}^n
\score_\theta(X_i)$ and information matrix $I_\theta =
\E_\theta[\score_\theta \score_\theta^\top]$. Moreover, in any family
$\{P_\theta\}$ that is LAN (Def.~\ref{definition:lan}) at $\theta$, if $T_n$
is a regular estimator (Def.~\ref{definition:regular-estimator}) at $\theta$
with limiting distribution $Q$, then for any bounded, symmetric,
quasi-convex loss $L$ and $c < \infty$,
\begin{equation*}
  \limsup_n \sup_{\norm{h} \le c}
  \E_{P_{\theta + h/\sqrt{n}}}
  \left[L(\sqrt{n}(T_n - \theta - h / \sqrt{n}))\right]
  = \E[L(W)] ~~ \mbox{for~}W \sim Q
\end{equation*}
(see Beran~\cite{beran1995role}, Eq.~(4.2)).
%
Thus, we show two results: first, that
that the family $\{P_\theta\}$ of distributions
defined by the shifted densities $\{f(\cdot - \theta)\}_{\theta \in \R}$
is quadratic-mean-differentiable at any $\theta$,
and second, that $\bar{\theta}_n$ is regular and
asymptotically normal.
The combination evidently gives the theorem.

First, we have the following lemma, somewhat more general than we need, but
whose proof we provide in Sec.~\ref{sec:proof-qmd}.
\begin{lemma}[Extension of \cite{VanDerVaart98}, Lemma 7.6]
  \label{lemma:qmd}
  Let $p_\theta$ be a density with respect to $\mu$,
  and assume that $\theta \mapsto s_\theta(x) \defeq \sqrt{p_\theta(x)}$
  is absolutely continuous for all $x$. Let
  $\dot{p}_\theta(x) = \nabla_\theta p_\theta(x)$ (when it exists),
  and assume that
  \begin{equation*}
    \mu(\{x : \dot{p}_\theta(x) ~ \mbox{fails~to~exist}\}) = 0.
  \end{equation*}
  Assume that $I_\theta \defeq \E_{P_\theta}[\dot{p}_\theta
    \dot{p}_\theta^\top / p_\theta^2]$ is continuous at $\theta_0$. Then
  $P_\theta$ is QMD (Definition~\ref{definition:qmd}) at $\theta = \theta_0$
  with $\score_\theta = \dot{p}_\theta / p_\theta$.
\end{lemma}

@@

This is essentially an immediate consequence of Le Cam's third
lemma~\cite[Example 6.7]{VanDerVaart98}.
Recall~\cite[Thm.~7.2]{VanDerVaart98} that if
a family $\{P_\theta\}_{\theta \in \Theta}$ is quadratic mean differentiable
at $\theta$ with score $\score_\theta$, then
it is LAN at $\theta$ (Definition~\ref{definition:lan})
with information matrix $I_\theta = \E[\score_\theta \score_\theta^\top]$.

%% then for any sequence
%% $h_n \to h$, if $Z_i \simiid P_\theta$ then
%% \begin{equation}
%%   \label{eqn:lan-from-qmd}
%%   \log \frac{dP_{\theta + h/\sqrt{n}}^n}{dP_\theta} (Z_1, \ldots, Z_n)
%%   = \frac{h}{\sqrt{n}} \sum_{i = 1}^n \score_\theta(Z_i)
%%   - \half h^2 I_\theta
%%   + o_{P_\theta,n}(1)
%% \end{equation}
%% where $I_\theta = \E_{P_\theta}[\score_\theta^2]$ is the Fisher information,
%% which necessarily exists.

%% We thus show that under the conditions we have assumed,
%% we have quadratic mean differentiability.

%% Deferring the proof of Lemma~\ref{lemma:qmd} temporarily,

The regularity result~\eqref{eq:normal_expansion_lem} gives
\begin{equation*}
  \sqrt{n} \bar{\Delta}_n = -\frac{1}{\sqrt{n}}
  \sum_{i = 1}^n  \frac{\varphi(Z_i)}{\psi'(0)} + o_{P,n}(1).
\end{equation*}
The conditions in
Corollary~\ref{corollary:normal-expansion}\eqref{item:apply-le-cam} imply
that the Fisher information $I_h =
\E_h[\score_h(Z)^2]$ exists and is continuous for
$\score_h(z) = \frac{p'(z - h)}{f(z - h)}$,
and the asymptotic expansion~\eqref{eqn:lan-from-qmd}
with the preceding display give the joint convergence
\begin{equation*}
  \left(\sqrt{n} \bar{\Delta}_n, \log \frac{P^n_{h_n/\sqrt{n}}}{P^n_0}(Z_1^n)
  \right)
  \cd \normal \left( \left(0,-\frac{h^2}{2} I_0 \right),
  \begin{pmatrix}
    \frac{\chi(0)}{\psi'(0)^2} & \frac{-h}{ \psi'(0)}
    \E_p[\varphi(Z) \score_0(Z)] \\
    \frac{-h}{ \psi'(0)} \E_p[\varphi(Z) \score_0(Z)]
    & h^2 I_0
  \end{pmatrix}  \right).
\end{equation*}
Le Cam's third lemma \cite[Exm. 6.7]{VanDerVaart98} then implies the
convergence
\begin{equation*}
  \sqrt{n}\bar{\Delta}_n
  \mathop{\cd}_{P_{h_n/\sqrt{n}}^n}
  \Ncal\left(\frac{-h}{ \psi'(0)} \E_f[\varphi(Z) \score(Z)],
  \frac{\chi(0)}{\psi'(0)^2} \right)
\end{equation*}
under the alternatives $P_{h_n/\sqrt{n}}$, which gives
Corollary~\ref{corollary:normal-expansion}\eqref{item:apply-le-cam}.

@@

\subsection{Proof of Lemma~\ref{lemma:qmd}}
\label{sec:proof-qmd}
